---
title: "Lab 6 Comparing two means"
author: "Tyler Edvall"
date: "04/22/21"
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

## Question A

> What is the mean difference in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

```{r}

# load packages -----------------------------------------------------------

library(tidyverse)


# read data ---------------------------------------------------------------

fish <- read_csv("chap12q19ElectricFish.csv")

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()


# data compiling ----------------------------------------------------------------

fish_long_summary <-
  fish_long %>% 
  group_by(location) %>% 
  summarize(
    n = n(),
    mean = mean(species),
    sd = sd(species),
    sem = sd/sqrt(n),
    upper = mean + 1.96 * sem,
    lower = mean - 1.96 * sem
  ) %>% 
  print()


t.test(formula = species ~ location, data = fish_long)

16.4-14.6

```

The mean difference is 1.8 and the 95% confidence interval of the mean is -4.587031 and 8.253697 between the number of species in the upstream versus the downstream areas of the tributaries. 


## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.


The p-value is 0.5596, which is greater than 0.05. This would indicate that we fail to reject the null hypothesis. 


## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.


We have to assume that the data is distributed normally and each of the two samples is a random sample from the population.

```{r}

fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("red","cyan4")) +
  theme_minimal()


fish_long %>% 
  ggplot(aes(x = location, y = species)) +
  geom_jitter(aes(color = location), 
              shape = 16, alpha = 1, width = 0.4) +
  geom_errorbar(aes(y = mean, ymax = upper, ymin = lower), 
                data = fish_long_summary, width = .1) +
  geom_point(aes(y = mean), 
             data = fish_long_summary) +
  scale_color_manual(values = c("red","cyan4")) +
  theme_minimal() +
  guides(color = "none")


```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

Graph the distribution of body temperatures for each crab type:

```{r}

library(readr)
crab <- read_csv("chap15q27FiddlerCrabFans.csv")

crab %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    bins = 8, 
    position = "identity",
    na.rm = TRUE
  ) + facet_wrap(~ crabType) +
  theme_minimal()

```

### Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

```{r}

aov_crab_temp_species <-
  aov(bodyTemperature ~ crabType, data = crab)

aov_crab_temp_species

summary(aov_crab_temp_species)

```

Null Hypothesis: The mean body temperature is equal among all crab species.

Alternative Hypothesis: At least one mean body temperature is different from the others.

The P-value is 7e-10, which is below any alpha level we could choose. Due to this, we would reject the null hypothesis. The conclusion that can be made is that the mean body temperature for at least one species is different from the means for the other species. 





















